How are galaxies assigned to halos? \\ Searching for assembly bias in the SDSS galaxy clustering

### 2016-10-06 text/ms.tex
final text prior to submission

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.
\title{How do galaxies get assigned to halos? \\ Searching for assembly bias in the SDSS galaxy clustering}

@@

Clustering of dark matter halos has been shown to depend on halo properties beyond mass such as halo concentration. This phenomenon is referred to as halo assembly bias. Standard halo occupation modeling (HOD) in large-scale structure studies assumes that halo mass alone is sufficient in characterizing the connection between galaxies and halos. Modeling of galaxy clustering can face systematic effects if the number or properties of galaxies are correlated with other halo properties. Using the Small MultiDark-Planck high resolution $N$-body simulation and the measurements of the projected two-point correlation function and the number density of Sloan Digital Sky Survey (SDSS) DR7 main galaxy sample, we investigate the extent to which the dependence of halo occupation on halo concentration can be constrained, and to what extent allowing for this dependence can improve our modeling of galaxy clustering.

Given the SDSS clustering data, our constraints on HOD with assembly bias, suggests that satellite population is not correlated with halo concentration at fixed halo mass. Furthermore, in terms of the number of centrals at fixed halo mass, our results favor lack of correlation with halo concentration in the most luminous samples ($M_{\rm r}<-21.5,-21$), modest levels of correlation for $M_{\rm r}<-20.5,-20, -19.5$ samples, lack of correlation for $M_{\rm r}<-19,-18.5$ samples, and anti-correlation for the faintest sample $M_{\rm r}<-18$.

We show that in comparison with abundance-matching mock catalogs, our findings suggest qualitatively similar but modest levels of assembly bias. Furthermore, by performing model comparison based on information criteria, we find that in most cases, the standard mass-only HOD model is still favored by the observations.

@@

HOD is an empirical framework that provides an analytic prescription for the expected number of galaxies $N$ that reside in halos by specifying a probability distribution function $P(N|x)$ where $x$ is a property of the halo. Under standard HOD modeling, the assumption is made that the halo mass $M$ alone is sufficient in determining the galaxy population of a halo. That is, the statistical properties of galaxies is governed by the halo mass. Mathematically, this assumption can be written as $P(N_g|M,\{x\})=P(N_g|M)$ where $\{x\}$ is the set of all possible halo properties beyond halo mass $M$.

Despite this very simplifying assumption, the models of galaxy--halo connection based on HOD modeling have been very successfully used in fitting the measurements of a wide range of statistics such as the projected two-point correlation function of galaxies, small scale redshift space distortion, three-point function, and galaxy--galaxy lensing with remarkable success (e.g. \citealt{zheng07,tinker_rsd2007,zehavi2011,leauthaud12,parejko2013,coupon2015,hod-3pcf,guo2015,miyatake15,zu2015,hod_vs_sham}). 
HOD has been used in constraining the cosmological parameters through modeling the galaxy two-point correlation function (hereafter 2PCF) (\citealt{abazajian2005}), combination of 2PCF with mass-to-light ratio of galaxies (\citealt{tinker05}), redshift space distortions (\citealt{tinker_rsd2007}), mass-to-number ratio of galaxy clusters (\citealt{tinker2012}) galaxy-galaxy weak lensing (\citealt{vdb03,cacciato13,more13,vdb13}) in the main sample of galaxies
of the Sloan Digital Sky Survey (hereafter SDSS, \citealt{york2000}), and also the combination of galaxy clustering and galaxy-galaxy lensing (\citealt{more15}) in SDSS III Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{boss}). Furthermore, HOD is implemented in producing mock galaxy catalogs in the BOSS survey (\citealt{manera2013,white2014}). It has also been used in galaxy evolution studies (\citealt{conroy2009,leauthaud12,behroozi13,hudson2015,zu2015,zu2016}).

This simple assumption of standard HODs  
does not sufficiently model the complexity of structure formation. Numerous $N$-body simulations that examine the clustering of 
dark matter halos have demonstrated that halo clustering is correlated with the formation 
history of halos. That is, at a fixed halo mass, the halo bias is correlated 
with properties of halos beyond mass, such as the concentration, formation time, or etc. 
This phenomenon is known as halo assembly bias (see \citealt{sheth2004,gao2005, harker2006, weschler2006, gao2007,croton2007,wang2007,angulo2008,dalal2008,li2008,sunayama2016}). 
There is even support for halo assembly bias in observations (\citealt{miyatake2016,more2016}).

Furthermore, the properties of galaxies within the halo may also depend on the halo's formation history. Then we may expect the occupation statistics of galaxies to be tied to the halo properties beyond mass such as the concentration of halos, environment, and etc. There have been many attempts in the literature at examining the dependence of galaxy properties on larger scale environment of the halos. But the results are mixed, and there is very little consensus. \citet{tinker_void_2009} show that the properties of the galaxies that reside in voids can be explained by the halo mass in which they live, and their properties are independent of their large scale environment of the halos. \citet{tinker_density_hod} proposes an extension of the standard HOD model $P(N|M)$ such that the number of galaxies residing in a halo not only depends on the mass of the halo, but also on the large scale density contrast $P(N|M,\delta)$. Based on modeling the clustering and void statistics of the SDSS galaxies, \citet{edHOD-tinker} shows that the dependence of the expected number of central galaxies on large scale density is not very strong.    
By randomly shuffling the galaxies among host halos of similar mass in the Millenium simulation, \citet{croton2007} show that assembly bias significantly impacts the galaxy two-point correlation functions. They also show that the effect is different for the faint and the bright samples. By applying a group finder to SDSS DR7 galaxies, \citet{tinker2011} shows that the quenched fraction of galaxies is independent of the large-scale environment.

Based on a group catalog of SDSS galaxies, \citet{blanton_group_2007} find that shuffling of galaxy colors among groups of similar mass has very little effect (only a slight change in small scales: less than $\sim 300 \rm{kpc}$) on the clustering of galaxies. Based on an alternative group catalog, \citet{wang2013} show that clustering of the central galaxies does not depend only on the halo mass, but also on additional parameters such as the specific star formation rate.

Another family of empirical galaxy--halo connection models is \emph{Abundance} \emph{Matching}. In Abundance Matching models, galaxies are assumed to live in halos/subhalos and are assigned luminosities or stellar masses by assuming a one-to-one mapping that preserves the rank ordering between its host halo/subhalo property and its luminosity or stellar mass
(\citealt{kravtsov2004,vale2004,tasitsiomi2004,conroy2009,guo2010,wetzel2010,Neisten2011,watson2012,rodriguez2012,kravstov2013,mao2015,chavez2016}). 
In practice, the standard abundance-matching models assume a monotonic mapping between the stellar mass/luminosity of galaxies and some host-halo property, such as 
the maximum circular velocity the host halo $V_\mathrm{max}$ that traces the depth of the gravitational potential well of the halo. Furthermore, a scatter is assumed in this one-to-one relation between galaxy property and halo/subhalo property. With this framework, abundance-matching models have been successfully used in modeling a wide range of the 
statistical properties of galaxies such as two-point correlation function (\citealt{reddick2013,lehman2015,hod_vs_sham}) as well as the group statistics of galaxies (\citealt{sham_gmf}).

It has been shown that the abundance-matching mock catalogs that use $V_{\rm max}$ (see \citealt{hw2013,arz2014}), or the ones that use some combination of $V_{\rm max}$ and host halo/subhalo virial mass $M_{\rm vir} $(see \citealt{lehman2015}) exhibit significant levels of assembly bias. That is, halo occupation in these models depends not only on halo mass, but also on other halo properties. This has been demonstrated by randomizing the galaxies among host halos of in bins of halo mass, such that the HOD remains constant, and comparing the difference in the 2PCF of the randomized catalog and the original mock catalog.

Abundance-Matching models can be extended to account for even higher orders of galaxy 
assembly bias. In so called \emph{Age-Matching} models, galaxy color is abundance-matched
to subhalo formation time so that older (redder) galaxies preferentially reside in older halos (\citealt{hw2013}). \citet{hearin2014} demonstrates that such an approach can reproduce a range of color and luminosity-dependent statistics of SDSS galaxies such as the 2PCF and galaxy-galaxy lensing. Based on the projected 2PCF measurements of (\citealt{hw2013}) galaxy catalogs, \citet{arz2014} showed that although the 2PCF measurements of these catalogs can be fit with the standard HOD modeling, the inferred HOD parameters cannot recover the underlying true HOD of these catalogs. That is, even in the presence of assembly bias in a galaxy sample, one can fit the clustering of this sample with HOD, but that does not guarantee recovery of the true HOD parameters. \citet{edHOD-weinberg} demonstrated that even with a galaxy catalog that has a significant level of assembly bias (e.g. \citealt{hw2013}), it is possible to extract the matter power spectrum more accurately if an environment-dependent HOD ($P(N|M,\delta)$) is used.

@@

Furthermore, we discuss how taking assembly bias into account in a more complex HOD model can improve our modeling of galaxy clustering in certain brightness limit. Then, we make a qualitative comparison between the levels of assembly bias in the mock catalogs created from our constrained on the decorated HOD model, and the \citet{hw2013} mock catalogs created using the abundance matching technique that uses maximum circular velocity $V_{\rm max}$ as subhalo property.

In order to investigate whether the additional complexity of the decorated HOD model is demanded by the galaxy clustering data, we perform a model comparison between the standard HOD model and the HOD model with assembly bias. We also discuss the effect of our choice of $N$-body simulation on our constraints, and previous works in the literature based on smaller $N$-body simulations (\citealt{zentner2016}).

@@

For our standard HOD model, we assume the HOD parameterization from \citet{zheng07}. 
According to this model, a dark matter halo can host a central galaxy and some number of satellite galaxies. The occupation of the central galaxies follows a nearest-integer distribution, 
and the occupation of the satellite galaxies follows a Poisson distribution. The expected number of centrals and satellites as a function of the host halo mass of $M_{h}$ are given by the following equations 
\begin{eqnarray}
\langle N_{\rm cen}(M_{h}) \rangle &=& \frac{1}{2}\Big[1+\Big(\frac{\log M_{h} - \log M_{\rm min}}{\sigma_{\log \rm{M}}} \Big) \Big], \label{hod:central}\\ 
\langle N_{\rm sat}(M_{h}) \rangle &=& \Big( \frac{M_{h} - M_{\rm{0}}}{M_{\rm 1}} \Big)^{\alpha}. \label{hod:satellite}
\end{eqnarray}

@@

In this occupation model, the allowable ranges that quantities $\Delta N_{c,i}$ and $\Delta N_{s,i}$ can take are given by 
\begin{eqnarray}
\mathrm{max} \{-\langle N_{c} | M_{h}\rangle, \langle N_{c} | M_{h}\rangle -1 \} \leq &\Delta N_{c,i}& \leq \mathrm{min} \{\langle N_{c} | M_{h}\rangle, 1-\langle N_{c} | M_{h}\rangle\}
 , \label{eq:cen-bounds} \\
-\langle N_{s} | M_{h}\rangle \leq & \Delta N_{s,i}& \leq \langle N_{s} | M_{h}\rangle. \label{eq:sat-bounds}
\end{eqnarray}
Afterwards, the assembly bias parameter $\mathcal{A}$ is defined in the following way:

@@

The advantage of using these measurements is that the effects of fiber collision systematic errors on the two-point statistics are corrected for (with the method described in \citealt{guo2012}), and therefore, these measurements provide accurate small scale clustering measurements. The assembly bias parameters introduced in section \ref{sec:method} can have a 10-percent level impacts on galaxy clustering (\citealt{decorated}). Presence of assembly bias in the satellite population impacts the very small-scale clustering (\citealt{decorated}). Moreover as \citet{sunayama2016} demonstrates, the scale-dependence of the halo assembly bias has a pronounced bump in the 1-halo to 2-halo transition regime (1$\sim$2 $h^{-1}\rm{Mpc}$). This scales can be impacted by fiber collision systematics. Precise investigation of the possible impact of this signal on the galaxy clustering modeling requires accurate measurements of 2PCF on small scales. 
The method of \citet{guo2012} is able to recover the true $\wpp$ with  $\sim 6\%$ accuracy in small scales ($r_{\rm{p}} = 0.1 \; h^{-1}\rm{Mpc}$) and with $\sim 2.5\%$ at relatively large scales $r_{\rm{p}} \sim 30 \; h^{-1}\rm{Mpc}$.

Note that the comoving volume of the $N$-body simulation used in this investigation is 64 $\times 10^{6} \; h^{-3}\rm{Mpc}^{3}$ which is larger than the comoving volume of all the luminosity-threshold samples in the SDSS data considered in this study except the two most luminous samples. The comoving volumes of the $M_{\rm r}<-21, \; 21.5$ samples are 71.74 and 134.65 (in units of $10^{6} h^{-3}\rm{Mpc}^{3}$) respectively. Since we are not studying very large scale clustering ($r_{\rm p} \leq 25 \; h^{-1}\rm{Mpc}$), using a slightly smaller box for those samples is justified.

@@

Figure \ref{fig:wpres} shows the fractional difference between the 68$\%$ and 95$\%$ posterior predictions of $w_p$ and the SDSS data. It is evident from Figure \ref{fig:wpres} that some improvement on modeling the clustering of the samples of $L_{\star}$ and slightly less brighter than $L_{\star}$ galaxies can be achieved by employing the more complex halo occupation model with assembly bias. As a result of apportioning more central galaxies to the high-$c$ halos relative to the low-$c$ halos, in the samples with luminosity thresholds of $M_{\rm r}<-20.5, -20, -19.5, -19$, the posterior predictions for $w_{p}$ are slightly improved in the intermediate scales ($1\sim 2 \; \mathrm{Mpc} h^{-1}$) and large scales. This can be also noted in significantly lower $\chi^{2}$ values---at the cost more model flexibility and higher degrees of freedom---achieved by the assembly bias model in these luminosity-threshold samples (see Table \ref{tab:constraints}).

@@

The satellite assembly bias can only alter the galaxy clustering at small-to-intermediate scales. Assigning more satellite galaxies to lower (or higher) concentration halos affects the one-halo term through increasing the satellite-satellite pair counts $\langle N_{\rm sat} N_{\rm sat} \rangle$. This results in boosting the small-scale clustering. But as pointed out by \citet{decorated}, the amount by which small-scale clustering increases also depends on the sign of the central assembly bias parameter $\acen$. Formation history of the halos can lead to the dependence of the abundance of subhalos on halo concentration (\citealt{zentner2005, mao2015}) at fixed halo mass, and since the occupation of the satellite galaxies is related to the abundance of subhalos, the satellite occupation may depend on halo concentration.

@@

It is important to note that the halo mass range in which the central assembly bias $\acen$ affects the central galaxy population is the mass range in which the condition $0<\langle N_{\rm cen}|M \rangle<1$ is met. Consequently, larger scatter parameter $\sigmam$ increases the dynamical mass range in which assembly bias affects the galaxy clustering. Note that in the luminosity regimes for which we obtain tighter constraints on $\acen$, the best-estimate values of the scatter parameter $\sigmam$ appear to be higher in the model with assembly bias. This is evident in Figure \ref{fig:posterior}. The model with assembly bias tends to push $\sigmam$ to higher values. This can be attributed to the tendency of this model to increase the effective dynamical mass range of central assembly bias.

@@

As shown in Table \ref{tab:constraints}, our constraints on the underlying standard HOD model obtained from the model with assembly bias and the model without assembly bias are in good agreement. The only cases in which there are mild tensions between the constraints found from the two models on the underlying HOD parameters, are the $M_{\rm r}<-20$ and the $M_{\rm r}<-20.5$ samples. However, these tensions are within one-sigma level. For instance, Figure \ref{fig:posterior} shows that in the $M_{\rm r}<-20.5$ sample, the constraint on the parameter $\alpha$ found from the model without assembly bias favor slightly higher values than the constraint found from the model with assembly bias. Also the scatter parameter $\sigmam$ is more tightly constrained in the standard HOD model. However, it is important to emphasize that these constraints are still in agreement with each other within a one-sigma level.

@@

For $L_{\star}$ galaxies, the difference in $w_{p}$ between the randomized and the original catalogs of \citet{hw2013} is shown in Figure \ref{fig:randomized} with the red curves. As demonstrated in Figure \ref{fig:randomized} (and as previously noted by \citealt{arz2014,edHOD-weinberg}), the relative difference in $w_{p}$ is only significant in large scales. This implies that in these catalogs, only the central galaxies are affected by assembly bias. This is in agreement with our findings.

Furthermore, we investigate whether the levels of assembly bias predicted by our findings are in agreement with the abundance matching catalogs of \citet{hw2013}. First, we make random draws from the posterior pdf over the parameters of the model with assembly bias. Then, we create mock catalogs with these random draws, and then we compute the difference in $w_{p}$ between the randomized catalogs and the original catalogs. The relative difference in $w_{p}$ predicted from our constraints are shown with blue curves in Figure \ref{fig:randomized}.

We note that our findings follow the same trend. That is, we see negligible difference in the small scale clustering and more considerable differences in $w_{p}$ on larger scales. Similar to findings of \citet{arz2014}, \citet{lehman2015}, and \citet{edHOD-weinberg}, we see that assembly bias becomes less significant in brighter galaxy samples. Furthermore, we notice that our mock catalogs favor more moderate levels of assembly bias as the fractional difference in $w_{p}$ is smaller in our mock catalogs.

@@

In terms of the effect of assembly bias on galaxy clustering, note that mocks created using the inferred parameters with the $\mathtt{SMDP}$ simulation show the same behavior as we observe in the abundance matching catalogs presented in \citet{arz2014} and \citet{lehman2015}. That is, the difference in $w_{p}$ between the mock catalogs and the randomized catalogs is mostly on large scales where the clustering is governed by the central galaxies.

>>>>

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.
\title{How are galaxies assigned to halos? \\ Searching for assembly bias in the SDSS galaxy clustering}

@@

Clustering of dark matter halos has been shown to depend on halo properties beyond mass such as halo concentration, a phenomenon referred to as halo assembly bias. Standard halo occupation modeling (HOD) in large scale structure studies assumes that halo mass alone is sufficient in characterizing the connection between galaxies and halos. Modeling of galaxy clustering can face systematic effects if the number or properties of galaxies are correlated with other halo properties. Using the Small MultiDark-Planck high resolution $N$-body simulation and the measurements of the projected two-point correlation function and the number density of Sloan Digital Sky Survey (SDSS) DR7 main galaxy sample, we investigate the extent to which the dependence of halo occupation on halo concentration can be constrained, and to what extent allowing for this dependence can improve our modeling of galaxy clustering.

Given the SDSS clustering data, our constraints on HOD with assembly bias, suggests that satellite population is not correlated with halo concentration at fixed halo mass. Furthermore, in terms of the occupation of centrals at fixed halo mass, our results favor lack of correlation with halo concentration in the most luminous samples ($M_{\rm r}<-21.5,-21$), modest levels of correlation for $M_{\rm r}<-20.5,-20, -19.5$ samples, lack of correlation for $M_{\rm r}<-19,-18.5$ samples, and anti-correlation for the faintest sample $M_{\rm r}<-18$.

We show that in comparison with abundance-matching mock catalogs, our findings suggest qualitatively similar but modest levels of assembly bias that is only present in the central occupation. Furthermore, by performing model comparison based on information criteria, we find that in most cases, the standard mass-only HOD model is still favored by the observations.

@@

HOD is an empirical framework that provides an analytic prescription for the expected number of galaxies $N$ that reside in halos by specifying a probability distribution function $P(N|x)$ where $x$ is a property of the halo. The assumption of the standard HOD modeling is that the halo mass $M$ alone is sufficient in determining the galaxy population of a halo. That is, the statistical properties of galaxies is governed by the halo mass. Mathematically, this assumption can be written as $P(N|M,\{x\})=P(N|M)$ where $\{x\}$ is the set of all possible halo properties beyond halo mass $M$.

Despite this simplifying assumption, the models of galaxy--halo connection based on HOD have been successfully used in fitting the measurements of a wide range of statistics such as the projected two-point correlation function of galaxies, small scale redshift space distortion, three-point function, and galaxy--galaxy lensing with remarkable success (e.g. \citealt{zheng07,tinker_rsd2007,zehavi2011,leauthaud12,parejko2013,coupon2015,hod-3pcf,guo2015,miyatake15,zu2015,hod_vs_sham}). 
HOD has been used in constraining the cosmological parameters through modeling the galaxy two-point correlation function (hereafter 2PCF) (\citealt{abazajian2005}), combination of 2PCF with mass-to-light ratio of galaxies (\citealt{tinker05}), redshift space distortions (\citealt{tinker_rsd2007}), mass-to-number ratio of galaxy clusters (\citealt{tinker2012}) galaxy-galaxy weak lensing (\citealt{vdb03,cacciato13,more13,vdb13}) in the main sample of galaxies
of the Sloan Digital Sky Survey (hereafter SDSS, \citealt{york2000}), and also the combination of galaxy clustering and galaxy-galaxy lensing (\citealt{more15}) in SDSS III Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{boss}). Furthermore, HOD is implemented in producing mock galaxy catalogs in the BOSS survey (\citealt{manera2013,white2014}). It has also been used in galaxy evolution studies (\citealt{conroy2009,leauthaud12,behroozi13,hudson2015,zu2015,zu2016}).

The complexity of structure formation however, is not sufficiently modeled under the standard HOD framework. Numerous $N$-body simulations that examine the clustering of 
dark matter halos have demonstrated that halo clustering is correlated with the formation 
history of halos. That is, at a fixed halo mass, the halo bias is correlated 
with properties of halos beyond mass, such as the concentration, formation time, or etc. 
This phenomenon is known as halo assembly bias (see \citealt{sheth2004,gao2005, harker2006, weschler2006, gao2007,croton2007,wang2007,angulo2008,dalal2008,li2008,sunayama2016}). 
There is support for halo assembly bias in observations of SDSS redMaPPer galaxy clusters (\citealt{miyatake2016}).

Furthermore, the properties of galaxies within the halo may also depend on the halo's formation history. Then we may expect the occupation statistics of galaxies to be tied to the halo properties beyond mass such as the concentration of halos. There have been many attempts in the literature at examining the dependence of galaxy properties on environment of the halos. But the results are mixed, and there is very little consensus. \citet{tinker_void_2009} show that the properties of the galaxies that reside in voids can be explained by the halo mass in which they live, and their properties are independent of their large scale environment of the halos. \citet{tinker_density_hod} proposes an extension of the standard HOD model $P(N|M)$ such that the number of galaxies residing in a halo not only depends on the mass of the halo, but also on the large scale density contrast $P(N|M,\delta)$. Based on modeling the clustering and void statistics of the SDSS galaxies, \citet{edHOD-tinker} shows that the dependence of the expected number of central galaxies on large scale density is not very strong.    
By randomly shuffling the galaxies among host halos of similar mass in the Millenium simulation, \citet{croton2007} shows that assembly bias significantly impacts the galaxy two-point correlation functions. They also show that the effect is different for the faint and the bright samples. By applying a group finder to SDSS DR7 galaxies, \citet{tinker2011} shows that the quenched fraction of galaxies is independent of the large-scale environment.

Based on a group catalog of SDSS galaxies, \citet{blanton_group_2007} find that shuffling of galaxy colors among groups of similar mass has very little effect (only a slight change in small scales: less than $\sim 300 \rm{kpc}$) on the clustering of galaxies. Based on an alternative group catalog, \citet{wang2013} shows that clustering of the central galaxies does not depend only on the halo mass, but also on additional parameters such as the specific star formation rate.

Another family of empirical galaxy--halo connection models is \emph{Abundance} \emph{Matching}. In Abundance Matching models, galaxies are assumed to live in halos and are assigned luminosities or stellar masses by assuming a monotonic mapping. In this monotonic mapping, the abundance of the halos are matched to the abundance of some property of galaxies (\citealt{kravtsov2004,vale2004,tasitsiomi2004,conroy2009,guo2010,wetzel2010,Neisten2011,watson2012,rodriguez2012,kravstov2013,mao2015,chavez2016}). 
One of the most commonly used host halo properties in abundance matching is the maximum circular velocity of the host halo $V_\mathrm{max}$ that traces the depth of the gravitational potential well of the halo. Furthermore, a scatter is assumed in this mapping. Within this galaxy--halo connection framework, abundance matching models have been successfully used in modeling a wide range of the statistical properties of galaxies such as two-point correlation function (\citealt{reddick2013,lehman2015,hod_vs_sham}) as well as the group statistics of galaxies (\citealt{sham_gmf}).

It has been shown that the abundance matching mock catalogs that use $V_{\rm max}$ (see \citealt{hw2013,arz2014}), or the ones that use some combination of $V_{\rm max}$ and host halo virial mass $M_{\rm vir}$ (see \citealt{lehman2015}) exhibit significant levels of assembly bias. That is, halo occupation in these models depends not only on halo mass, but also on other halo properties. This has been demonstrated by randomizing the galaxies among host halos in bins of halo mass, such that the HOD remains constant, and then comparing the difference in the 2PCF of the randomized catalog and that of the original mock catalog.

Age Matching models extend the Abundance Matching models to account for even higher orders of galaxy 
assembly bias. In these models, galaxy color is matched to halo formation time so that older (redder) galaxies preferentially reside in older halos (\citealt{hw2013}). \citet{hearin2014} demonstrates that such an approach can reproduce a range of color and luminosity-dependent statistics of SDSS galaxies such as the 2PCF and galaxy-galaxy lensing. Based on the projected 2PCF measurements of (\citealt{hw2013}) galaxy catalogs, \citet{arz2014} showed that although the 2PCF measurements of these catalogs can be fit with the standard HOD modeling, the inferred HOD parameters cannot recover the underlying true HOD of these catalogs. That is, even in the presence of assembly bias in a galaxy sample, one can fit the clustering of this sample with HOD, but that does not guarantee recovery of the true HOD parameters. \citet{edHOD-weinberg} demonstrated that even with a galaxy catalog that has a significant level of assembly bias (e.g. \citealt{hw2013}), it is possible to extract the matter power spectrum more accurately if an environment-dependent HOD ($P(N|M,\delta)$) is used.

@@

Furthermore, we discuss how taking assembly bias into account in a more complex HOD model can improve our modeling of galaxy clustering in certain brightness limits. Then, we make a qualitative comparison between the levels of assembly bias in the mock catalogs created from our constraints on the decorated HOD model, and the \citealt{hw2013} mock catalogs created using the abundance matching technique that uses maximum circular velocity $V_{\rm max}$ as subhalo property. Our comparison shows the levels of assembly bias seen in the predictions of both models follow the same trend. That is assembly bias is more prominent in lower luminosity-threshold samples and its impact on galaxy clustering is only significant on large scales.

In order to investigate whether the additional complexity of the decorated HOD model is demanded by the galaxy clustering data, we perform a model comparison between the standard HOD model and the HOD model with assembly bias. We also discuss the effect of our choice of $N$-body simulation on our constraints, and previous works in the literature (\citealt{zentner2016}) based on smaller $N$-body simulations. In addition to analysis of the luminosity-threshold samples presented in \citet{zentner2016}, we consider the faintest ($M_{\rm r} < -18,-18.5$) and the brightest ($M_{\rm r}<-20.5$) galaxy samples. For the samples considered in both \citet{zentner2016} and this investigation, we compare the constraints on the expected levels of assembly bias.

@@

For our standard HOD model, we assume the HOD parameterization from \citet{zheng07}. 
According to this model, a dark matter halo can host a central galaxy and some number of satellite galaxies. The occupation of the central galaxies follows a nearest-integer distribution, 
and the occupation of the satellite galaxies follows a Poisson distribution. The expected number of centrals and satellites as a function of the host halo mass of $M_{h}$ are given by the following equations 
\begin{eqnarray}
\langle N_{\rm c}| M_{h} \rangle &=& \frac{1}{2}\Big[1+\Big(\frac{\log M_{h} - \log M_{\rm min}}{\sigma_{\log \rm{M}}} \Big) \Big], \label{hod:central}\\ 
\langle N_{\rm s} | M_{h} \rangle &=& \Big( \frac{M_{h} - M_{\rm{0}}}{M_{\rm 1}} \Big)^{\alpha}. \label{hod:satellite}
\end{eqnarray}

@@

In this occupation model, the allowable ranges that quantities $\Delta N_{c,1}$ and $\Delta N_{s,1}$ can take are given by 
\begin{eqnarray}
\mathrm{max} \{-\langle N_{c} | M_{h}\rangle, \langle N_{c} | M_{h}\rangle -1 \} \leq &\Delta N_{c,i}& \leq \mathrm{min} \{\langle N_{c} | M_{h}\rangle, 1-\langle N_{c} | M_{h}\rangle\}
 , \label{eq:cen-bounds} \\
-\langle N_{s} | M_{h}\rangle \leq & \Delta N_{s,i}& \leq \langle N_{s} | M_{h}\rangle. \label{eq:sat-bounds}
\end{eqnarray}
Afterwards, the assembly bias parameter $\mathcal{A}$ is defined in the following way:

@@

The advantage of using these measurements is that the effects of fiber collision systematic errors on the two-point statistics are corrected for (with the method described in \citealt{guo2012}), and therefore, these measurements provide accurate small scale clustering measurements. The assembly bias parameters introduced in section \ref{sec:method} can have a 10-percent level impacts on galaxy clustering (\citealt{decorated}). Presence of assembly bias in the satellite population impacts the very small-scale clustering (\citealt{decorated}). Moreover as \citet{sunayama2016} demonstrates, the scale-dependence of the halo assembly bias has a pronounced bump in the 1-halo to 2-halo transition regime (1$\sim$2 $h^{-1}\rm{Mpc}$). This scale can be impacted by fiber collision systematics. Precise investigation of the possible impact of this signal on the galaxy clustering modeling requires accurate measurements of 2PCF on small scales. 
The method of \citet{guo2012} is able to recover the true $\wpp$ with  $\sim 6\%$ accuracy in small scales ($r_{\rm{p}} = 0.1 \; h^{-1}\rm{Mpc}$) and with $\sim 2.5\%$ at relatively large scales $r_{\rm{p}} \sim 30 \; h^{-1}\rm{Mpc}$.

Note that the comoving volume of the $N$-body simulation used in this investigation is 64 $\times 10^{6} \; h^{-3}\rm{Mpc}^{3}$ which is larger than the comoving volume of all the luminosity-threshold samples in the SDSS data considered in this study except the two most luminous samples. The comoving volumes of the $M_{\rm r}<-21, \; 21.5$ samples are 71.74 and 134.65 (in units of $10^{6} h^{-3}\rm{Mpc}^{3}$) respectively. Since we are not studying very large scale clustering ($r_{p,\; \rm max} \leq 25 \; h^{-1}\rm{Mpc}$), using a slightly smaller box for those samples is justified.

@@

Figure \ref{fig:wpres} shows the fractional difference between the 68$\%$ and 95$\%$ posterior predictions of $w_p$ and the SDSS data. It is evident from Figure \ref{fig:wpres} that some improvement on modeling the clustering of the samples of $L_{\star}$ and slightly less brighter than $L_{\star}$ galaxies can be achieved by employing the more complex halo occupation model with assembly bias. As a result of apportioning more central galaxies to the high-$c$ halos relative to the low-$c$ halos, in the samples with luminosity thresholds of $M_{\rm r}<-20.5, -20, -19.5$, the posterior predictions for $w_{p}$ are slightly improved in the intermediate scales ($1\sim 2 \; \mathrm{Mpc} h^{-1}$) and large scales. This can be also noted in significantly lower $\chi^{2}$ values---at the cost more model flexibility and higher degrees of freedom---achieved by the assembly bias model in these luminosity-threshold samples (see Table \ref{tab:constraints}).

@@

The satellite assembly bias can only significantly alter the galaxy clustering at small-to-intermediate scales. Assigning more satellite galaxies to lower (or higher) concentration halos affects the one-halo term through increasing the satellite-satellite pair counts $\langle N_{\rm s} N_{\rm s} \rangle$. This results in boosting the small-scale clustering. But as pointed out by \citet{decorated}, the amount by which small-scale clustering increases also depends on the sign of the central assembly bias parameter $\acen$. Formation history of the halos can lead to the dependence of the abundance of subhalos on halo concentration (\citealt{zentner2005, mao2015}) at fixed halo mass, and since the occupation of the satellite galaxies is related to the abundance of subhalos, the satellite occupation may depend on halo concentration.

@@

It is important to note that the halo mass range in which the central assembly bias $\acen$ affects the central galaxy population is the mass range in which the condition $0<\langle N_{\rm c}|M \rangle<1$ is met. Consequently, larger scatter parameter $\sigmam$ increases the dynamical mass range in which assembly bias affects the galaxy clustering. Note that in the luminosity regimes for which we obtain tighter constraints on $\acen$, the best-estimate values of the scatter parameter $\sigmam$ appear to be higher in the model with assembly bias. This is evident in Figure \ref{fig:posterior}. The model with assembly bias tends to push $\sigmam$ to higher values. This can be attributed to the tendency of this model to increase the effective dynamical mass range of central assembly bias.

@@

As shown in Table \ref{tab:constraints}, our constraints on the underlying standard HOD model obtained from the model with assembly bias and the model without assembly bias are in good agreement. The only cases in which there are mild tensions between the constraints found from the two models on the underlying HOD parameters, are the $M_{\rm r}<-20$ and the $M_{\rm r}<-20.5$ samples. However, these tensions are still within one-sigma level. For instance, Figure \ref{fig:posterior} shows that in the $M_{\rm r}<-20.5$ sample, the constraint on the parameter $\alpha$ found from the model without assembly bias favor slightly higher values than the constraint found from the model with assembly bias. Also the scatter parameter $\sigmam$ is more tightly constrained in the standard HOD model. However, it is important to emphasize that these constraints are still in agreement with each other within a one-sigma level.

@@

For $L_{\star}$ galaxies, the difference in $w_{p}$ between the randomized and the original catalogs of \citet{hw2013} is shown in Figure \ref{fig:randomized} with the red curves. As demonstrated in Figure \ref{fig:randomized} (and as previously noted by \citealt{arz2014,edHOD-weinberg}), the relative difference in $w_{p}$ is only significant in relatively large scales. This implies that in these catalogs, only the central occupation are affected by assembly bias. This is in agreement with our findings.

Furthermore, we investigate whether the levels of assembly bias predicted by our findings are in agreement with the abundance matching catalogs of \citet{hw2013}. First, we make random draws from the posterior probability distribution function over the parameters of the model with assembly bias. Then, we create mock catalogs with these random draws, and then we compute the difference in $w_{p}$ between the randomized catalogs and the original catalogs. The relative difference in $w_{p}$ predicted from our constraints are shown with blue curves in Figure \ref{fig:randomized}.

We note that our findings follow the same trend. That is, we see negligible difference in the small scale clustering and more considerable differences in $w_{p}$ on larger scales ($r_{p} > 1 Mpc \; h^{-1}$). Similar to findings of \citet{arz2014}, \citet{lehman2015}, and \citet{edHOD-weinberg}, we see that assembly bias becomes less significant in brighter galaxy samples. Furthermore, we notice that our mock catalogs favor more moderate levels of assembly bias as the fractional difference in $w_{p}$ is smaller in our mock catalogs.

@@

In terms of the effect of assembly bias on galaxy clustering, note that mocks created using the inferred parameters with the $\mathtt{SMDP}$ simulation show the same behavior as we observe in the abundance matching catalogs presented in \citet{arz2014} and \citet{lehman2015}. That is, the difference in $w_{p}$ between the mock catalogs and the randomized catalogs is mostly on large scales where the clustering is governed by the central galaxies. That is, the impact of assembly bias on the satellite occupation is negligible and only the central occupation is affected.
